Fisheries management when quotas are costly to change

### 2014-08-21 manuscripts/crashes.Rmd
additional analyses
updated presentation, manuscript references, added quad_costs to appendix, with relevant code

library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

```{r analysis, include=FALSE}
source("components/analysis.R")
```

>>>>

library("methods")
library("knitr")
basename <- "crashes"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

@@

```{r analysis, include=FALSE}
source("components/crash-analysis.R")
```

### 2014-08-21 manuscripts/manuscript.Rmd
updated presentation, manuscript references, added quad_costs to appendix, with relevant code

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTK
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK

address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: UCD
    address: Department of Environmental Science and Policy, University of California, Davis
  - code: melbourne
    address: University of Melbourne, Australia 
  - code: UTK
    address: University of Tennessee, Knoxville

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  - to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. While some types of cost act to smooth out variation in
          quotas and stock sizes, as one might have expected, others can actually
          increase variation in stock sizes and quotas through time. We also show
          that the potential economic impact on the fishery of managers assuming
          policy adjustment costs are present when in fact they are absent is much
          smaller than the impact of managers assuming such costs are absent if
          they are present.

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

```{r Figure_1, fig.cap="ICCAT Stock assessments and US Quotas for Bluefin Tuna, 1987 - 2007", fig.width=6, fig.height=4, dev=c("pdf", "svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
tuna<-melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna harvest (tonnes)") + theme_bw()
```

To illustrate these concepts, Figure 1 show an example from fisheries
management. On the left axis, the figure shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  On the right axis the figure
also shows the catch quota for the stock that was set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, which could contribute to
the observed quota stability. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

@@

With his formulation, Reed showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. But any time there is a good recruitment pulse, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

@@

\begin{equation}
\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^\infty \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}
\end{equation}

where $\mathbf{E}$ is the expectation operator, $\delta>0$ is the
discount rate and $\Pi_0$ is the net revenue from operating the fishery
in a given year. In this base case, we assume the annual net revenue
from fishing is

@@

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$. (Examples
shown here use $p$= `r price` and $c_0$ = `r c0`, complete code is
provided in the supplementary materials to re-run these analyses with
arbitrarily specified values and replot the figures shown).

@@

_Economic penalties promised for the supplement are still to come_

```{r Figure_S1, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r Figure_S2, dependson="tidy", fig.cap="Example realization of resulting fish stock sizes under the different functional forms of adjustment costs, corresponding to policies shown in Figure 3.  Because stock sizes are measured before harvests, the effect of the penalty on the stock size is not as pronounced.", fig.height=3, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

>>>>

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTK_Econ
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK_EEB

address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: UCD
    address: Department of Environmental Science and Policy, University of California, Davis
  - code: melbourne
    address: School of Botany, University of Melbourne, Australia 
  - code: UTK_EEB
    address: Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville
  - code: UTK_Econ
    address: Department of Economics, University of Tennessee, Knoxville

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  - to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. While some types of cost act to smooth out variation in
          quotas and stock sizes, as one might have expected, others can actually
          increase variation in stock sizes and quotas through time. We also show
          that the potential economic impact on the fishery of managers assuming
          policy adjustment costs are present when in fact they are absent is much
          smaller than the impact of managers assuming such costs are absent if
          they are present. Further, we demonstrate that the structural form of
          these costs matters -- assuming one form when costs are are in fact
          a different form can be even more costly.

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space [@Durrett1994]. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

```{r Figure_1, fig.cap="ICCAT Stock assessments and Total Allowable Catch (TAC, before adjustments for national underages and overages) for Bluefin Tuna, 1987 - 2007", fig.width=6, fig.height=4, dev=c("pdf", "svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
tuna <- melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna harvest (tonnes)") + theme_bw()
```

To illustrate these concepts, we use an example from fisheries
management. Fig1a shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT20XX].  Fig1b 
shows the catch quota (Total Allowable Catch or TAC, before being adjusted
for overages or underages in the catch of each country in the previous year)
 for the stock as set by the relevant
management agency [@ICCAT20XX]. Despite the estimated population size
declining by XX% between YYYY and ZZZZ, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, which could contribute to
the observed quota stability. More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

@@

With his formulation, @Reed1979 showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. But any time there is a good recruitment pulse, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

@@

\begin{equation}
\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^{T_{\mathrm{max}}} \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}
\end{equation}

where $\mathbf{E}$ is the expectation operator, $\delta>0$ is the
discount rate and $\Pi_0$ is the net revenue from operating the fishery
in a given year. In this base case, we assume the annual dockside (that is, before accounting for any adjustment costs that may be incurred) net revenue
from fishing is

@@

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$. (Examples
shown here use $p$= `r price` and $c_0$ = `r c0`.  Sensitivity analysis appears
in the supplement, along with complete R code to replicate and further 
vary the setup).

@@

## Economic penalties using quadratic costs

```{r quadraticcosts}
source("components/quad_costs.R")
```

Calibrated in the same way.

```{r Figure_S1}
ex <- dt_quad %>% 
  filter(replicate=="rep_17" & time < 20) %>%
  select(time, harvest, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=c(harvest = substitute(paste("Harvest Policy under ", paste(Pi[4]))), 
                                harvest_alt = substitute(paste("Harvest under ", paste(Pi[0]), " (Reed)"))), 
                       name="")

```{r}
shifts <- dt[, fraction_no_shift(harvest), by=c("replicate")]
ave <- shifts[, mean(V1)]
```

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of contant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

```{r Figure_S2, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r Figure_S3, dependson="tidy", fig.cap="Example realization of resulting fish stock sizes under the different functional forms of adjustment costs, corresponding to policies shown in Figure 3.  Because stock sizes are measured before harvests, the effect of the penalty on the stock size is not as pronounced.", fig.height=3, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

### 2014-08-21 manuscripts/presentation.Rmd
updated presentation, manuscript references, added quad_costs to appendix, with relevant code

@@

@@

```{r, fig.cap="",  dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent"), fig.height=3}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "L1" & time < 20) %>%
  select(time, fishstock, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=c(fishstock = "Stock size", harvest_alt = "Harvest policy"), name="")

>>>>

## Progress in modeling ecological complexity:

- Importance of space
- temporal heterogeneity 
- individual heterogeneity
- ecosystem models

@@

<blockquote class="twitter-tweet" lang="en"><p>No public policy issue is ultimately about science; it’s about stakeholders, values, economics, ideology. Goldston <a href="https://twitter.com/hashtag/ESA2014?src=hash">#ESA2014</a></p>&mdash; Liza Gross (@lizabio) <a href="https://twitter.com/lizabio/statuses/498971982998687744">August 11, 2014</a></blockquote>
<script async src="widgets.js" charset="utf-8"></script>

## Hmm...

## Can we put that in our models too?

@@

```{r, fig.cap="",  dev=c("svg", "png"), cache=FALSE, dev.args=list(bg="transparent"), fig.height=3}
ex <- dt %>% 
  filter(replicate=="rep_17" & penalty_fn == "L1" & time < 20) %>%
  select(time, fishstock, harvest, alternate, harvest_alt) %>%
  gather(variable, stock, -time) 
ggplot(ex) + 
  geom_line(aes(time, stock, col=variable), lwd=1) + 
  scale_color_discrete(labels=c(fishstock = "Stock Size", harvest = "Harvest", alternate = "Stock size (Adj)", harvest_alt = "Harvest policy (Adj)"), name="")
